---
title: "Homework week 39"
author: "Caroline"
date: " created 4 October 2021 by Adela Sobtkova, updated 3 October 2023 by Caroline"
output: 
html_document:
  toc:true
  toc_float:true
---
## 1) Visualize the duration of reign among the danish kings 
I downloaded my danish kings dataset by using read_csv

```{r warning=FALSE}
library(tidyverse)
library(ggplot2)

read_csv("data/Tidy_Data_W37_CLEAN_danish_kings.csv")
danish_kings <- read_csv("data/Tidy_Data_W37_CLEAN_danish_kings.csv")

danish_kings %>% 
  ggplot(aes(x=Start_of_Reign,y=Reign_Sum))+
  geom_col(color="red")+
  geom_smooth()+
  labs(x="time",y="duration_of_reign")

```

We already had a column in our spreadsheet named Reign_Sum, which contains the data of how long each king reigned.
We plotted our data by using ggplot, where we decided to use the geom_point.
We assigned Start_of_reign as our x-axis to create "a timeline", and we assigned Reign_sum as our y-axis.
We created a trend line by adding geom_smooth, which shows us the diachronic trend in the duration of reign from ca. 900-today


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE)
```
This exercise is based on the dataset provided by [OurWorldInData project](https://ourworldindata.org/homicides) based at the Oxford University.

## 2) The long-term trend in Homicides in Western Europe

Understanding how homicide rates have changed prior to the modern era requires the help of historians and archivists. Manuel Eisner, a criminology professor at the University of Cambridge, and his colleagues published the Historical Violence Database \href{https://cjrc.osu.edu/research/interdisciplinary/hvd}: a compilation of data on long-term trends in homicide rates, in addition to qualitative information such as the cause of death, perpetrator and victim. This database is limited to countries with relatively complete historical records on violence and crime – mainly Western Europe and the US.

Starting in the second half of the nineteenth century, these European regions have consistent police records of those accused of murder or manslaughter and annual counts of homicide victims. To go back further in time, reaching as far back as the thirteenth century, Eisner collected estimates (from historical records of coroner reports, court trials, and the police) of homicide rates made in over ninety publications by scholars.

Homicide rates – measured as the number of homicides per 100,000 individuals – up to 1990 are sourced from Eisner’s (2003) publication and the Historical Violence Database. 

Are homicide rates in Europe today lower or higher than in the past? Using the provided dataset, display and describe the long-run homicide rates for the five European regions: Italy, England, Germany, Netherlands and Scandinavia.

```{r library}
library(tidyverse)
```

## Load the available data from ourworldindata.org

You should always interrogate the source of your data. Who compiled it, from where, what is missing, how representative the data are? Check the data/Metadata.txt to learn about the data provenance.

```{r load-data}
Western_Europe <- read_csv("data/homicide-rates-across-western-europe.csv")
```

## Inspect the data

How clean and analysis-ready is the dataset? Do you understand what the column names represent? What is the difference between rate and homicide number?

```{r inspect}
head(Western_Europe)
```

Ok, the data look good except for the column `Homicide rate in Europe over long-term (per 100,000)` which is not very easy to work with.

- Use the `names()` function and assignment key to relabel this column to `homicides_per_100k` 

```{r relabel column}
names(Western_Europe) <- c("entity","code","year","homicide_per_100k")
head(Western_Europe)
```

Now, that you have looked at what the data looks like and what it represents, and streamlined it, let's see what big picture it contains.

## Let's see what the long-term trend is in homicides

- use `ggplot()` function and remember the`+` at the end of the line
- chose a meaningful `geom_......()`  for geometry (hint: points are not great)
- load `Year` on the `x` axis and `homicides_per_100k` column in y axis
- to color individual country entries consistently, assign the country column to the argument `color`.
- provide meaningful title and axis labels
- remember to change the `eval` flag so that the code chunk renders when knitted 

```{r plot-homicides}
ggplot(data = Western_Europe)+
  geom_col(aes(x=year,y=homicide_per_100k,color=entity))+
  labs(title = "Homicide rates in Western Europe")+
  labs(x="Year",y="Number of homcides per 100,000 people")+
  labs(color="Countries")

```
Alright, the homicide rates should all be descending over time. What a comfort. But the viz is not super clear. Let's check the rates for individual countries.

## Uncouple the homicides of individual countries for easier view

You can visualize each country's trend separately by adding an extra argument to the ggplot, the `facet_wrap()` and feeding it the country column. If in doubt, check your ggplot tutorial and your country column name for exact usage.

- reuse the ggplot from the chunk above
- insert `facet_wrap()` after the specification of geometry to split countries in separate charts
- change the facet "layout" to two columns and three rows so that the trends are easier to see in horizontal layout. 


```{r eval = FALSE}
ggplot(data = Western_Europe)+
  geom_col(aes(x=year,y=homicide_per_100k,color=entity))+
  facet_wrap(~entity,ncol=2)+
  labs(title = "Homicide rates in Western Europe")+
  labs(x="Year",y="Number of homcides per 100,000 people")+
  labs(color="Country")+
  theme(legend.position = "bottom")

```

## Compare the trends in homicide with the pattern of reign duration among Danish rulers through time.

- Load your Danish king dataset. Hopefully it is tidy and your years and duration of reign are all numeric.
- You need to have a consistent way of plotting the rulers' reign on the x axis, so I recommend you create a midyear column by calculating the middle of each monarch's rule (Hint: `midyear` = `endyear` - (`endyear`-`startyear`)/2)
- Start a ggplot plotting midyear on x axis and duration on y axis
- Try `geom_smooth()` for geometry 
- Provide meaningful labels and a title
- How would you characterize the trend compared to the homicides above?
```{r}
danish_kings %>% 
  mutate(Midyear=End_of_reign-(End_of_reign-Start_of_Reign)/2) %>% 
  ggplot(aes(x=Midyear,y=Reign_Sum))+
  geom_col(color="red")+
  geom_smooth()+
  labs(x="time",y="duration of reign")
  
```
First of all our danish_kings dataset was already downloaded because we used it in task 1

The homicide rate is generally decreasing through time. The homicide rate has been very low in Scandinavia since the mid 1700th century, which is the same time as the danish kings reigned for the longest time. The trend in the duration of reign among the danish kings is slightly increasing through time and is at its highest around year 1650 and the kings continues to reign for a long peiode of time. Therefore there might be a connection between the consitency of a regent and the homicide rate in a country. It is not a very clear connection though.  

#Final tasks:

1) Plot: In the facetted plot above, move the legend from the current position on the side to below the facets, and label it "Country" instead of "Entity".

```{r}
ggplot(data = Western_Europe)+
  geom_col(aes(x=year,y=homicide_per_100k,color=entity))+
  facet_wrap(~entity,ncol=2)+
  labs(title = "Homicide rates in Western Europe")+
  labs(x="Year",y="Number of homcides per 100,000 people")+
  labs(color="Country")+
  theme(legend.position = "bottom")
```
Like we did in the previous task:)

2) Rmarkdown: 

  - edit the author of the document, and convert 'Final Tasks' into heading #2 (like the other headings) 
  - add a `floating table of contents` to your Rmarkdown document, 
  - provide informative `chunk-names` and edit flags in your R chunks, and
  - automatically generate a `timestamp` to show when the document was last updated. (Hint: check the Rmarkdown episode in our Data Carpentry tutorial)

3) Question: In <250 words articulate your answer on the basis of the data visualisations to the following question: are we more civilized today?
When only having the data on the homicide rates in Western Europe it is hard to conclude whether we are more civilized today. Looking at the the homicide rates it is sure to say that the amount of homicides comitted in Europe for the past 1000 years has been decreasing, especially from around the mid 17th century till today. Not killing eachother indicates stability, which is a result of political, economic and legal structures developed in the western societies through time. It can be argued that these societal structures are a sign of more civilized societies. 
